---
title: "Relationship between Titratable Acidity, pH, and Taste Rating"
author: "Graciela Ferderer"
date: "2023-08-16"
output: github_document
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## Evaluation of Titratable Acidity as a Quality Metric

Titratable Acidty (TA) is the approximation of the amount of weak organic acids in a solution. Kombucha produces a variety of these acids, primarily acetic acid. In effort to improve our quality metrics we seek evidence to determine TA as a good indicator of the target flavor profile. 

This analysis investigates the hypothesis that high TA is a good indicator of a high Taste Rating. 

## Data Sources

Data was sourced from our Production Brewing Database. This database is maintained by the Quality Manager and is updated weekly from measurements recorded in our lab. Data used was collected from January to July 2023.

```{r library calls, echo = FALSE, message = FALSE}

installed.packages("ggplot2", quiet = TRUE)
library(ggplot2)

options(repos=c(CRAN="https://cloud.r-project.org/"))
install.packages("ggpubr", quiet = TRUE)
library(ggpubr)

brew_data <- read.csv("C:\\Users\\HMKom\\Downloads\\Production_Brewing_Data - Cleaned Production Brewing Data.csv", header=TRUE, stringsAsFactors=FALSE) #loading brew_data

brew_data$Final.Titratable.Acidity <- as.numeric(brew_data$Final.Titratable.Acidity) #converting the data type of column Final.Titratable.Acidity

```
## Data Cleaning Documentation in Google Sheets

1. **Remove duplicates:** To ensure data integrity, remove duplicate entries from the dataset. Navigate to "Data" > "Data Clean up" > "Remove Duplicates".

2. **Delete extraneous columns:** Remove any unnecessary columns, such as helper columns or target values, that are not relevant to the analysis.

3. **Delete extraneous blank cells at the bottom:** Remove any empty rows at the bottom of the dataset that do not contain any useful information.

4. **Format columns into corresponding integer type:** Select the relevant columns and format them as integers to ensure consistent data types. To do this, select the whole column, go to "Format" > "Select Data Type".

5. **Filter out null values:** Remove rows with null or missing values from the dataset to maintain data quality. Use the "Data" > "Add Filter" option to filter out null values.

6. **Split helper columns and extract starter birth date:** If there are helper columns that contain additional information, split those columns and extract the starter birth date using an appropriate technique.

7. **Duplicate Starter Column:** Create a duplicate of the starter column for further modifications.

8. **Add two columns to the right of the duplicated column:** Insert two new columns next to the duplicated starter column for additional data processing.

9. **Split text to columns using the period separator:** Use the "Data" > "Split Text to Columns" function to split the data in the duplicated column into separate columns based on the period separator. This will separate the data into multiple columns.

10. **Label subsequent columns:** Assign appropriate labels to the newly created columns, such as "Creation Date" and "Parent Starter", to indicate the information they represent.

11. **Delete duplicate starter column:** Remove the duplicated starter column now that the required information has been extracted and processed.



## Summary of Analysis in R

1. **Structure of the Data Frame:** The str() function was used to view the structure of the data frame. It provides information about the variables and their data types in the "brew_data" data frame. This step helps in identifying any data types that may need to be converted for further analysis.

2. **Conversion of Data Type:** The data type of the "Final.Titratable.Acidity" column was converted using the as.numeric() function. The column was reassigned the converted numeric values using the assignment operator (<-). This conversion ensures that the "Final.Titratable.Acidity" column is treated as numeric data for subsequent analysis.

3. **Generation of Scatter Plots:** Scatter plots with linear trend lines and the Spearman correlation coefficient were generated using the ggplot() function from the ggplot2 package. The geom_point() function was used to add the scatter plot points to the plot, while the geom_smooth(method = lm) function added the linear trend line. The stat_cor(method = "spearman") function was utilized to display the Spearman correlation coefficient label on the plot.

By examining the structure of the data frame, converting the data type of the "Final.Titratable.Acidity" column to numeric, and generating scatter plots with linear trend lines and the Spearman correlation coefficient, the analysis ensures proper understanding of the data and enables further exploration of the relationships between variables.

## Titratable Acidity and Taste Rating have a Moderate Correlation
```{r TA v Taste, echo=FALSE}
#Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
#Relationship between Titratable Acidity and Taste Rating
ggplot(brew_data, aes(x=Final.Titratable.Acidity, y=Taste.Rating.out.of.10))+
  geom_point()+
  geom_smooth(method = lm )+ #adds linear trend line
  stat_cor(method = "spearman", label.x = 4, label.y = 6.5) #adds spearman coefficient
```

The p value \<1% indicates these results are statistically significant. The R value of 0.56 indicates these is a moderate correlation between increase in TA and taste rating.

## pH and Taste Rating have a Moderate Correlation
```{r Final pH v Taste, echo=FALSE}
ggplot(brew_data, aes(x=Final.pH, y=Taste.Rating.out.of.10))+
  geom_point()+
  geom_smooth(method = lm )+ #adds linear trend line
  stat_cor(method = "spearman", label.x = 2.9, label.y = 6) #adds spearman coefficient
```

The p value \<1% indicates these results are statistically significant. The R value of -0.49 indicates these is a moderate correlation between decrease in pH and taste rating.

## Titratable Acidity and pH have a Strong Correlation
```{r Taste v pH, echo=FALSE}
#Relationship between Final pH and Titratable Acidity
ggplot(brew_data, aes(x=Final.Titratable.Acidity, y=Final.pH))+
  geom_point()+
  geom_smooth(method = lm )+ #adds linear trend line
  stat_cor(method = "spearman", label.x = 2, label.y = 3) #adds spearman coefficient
```

The p value \<1% indicates these results are statistically significant. The R value of -0.72 indicates these is a moderate correlation between decrease in pH and increase in TA.

## Discusion 

pH and Titratable Acidity (TA) both have a moderate correlation with Taste Rating. In comparing R Values TA and Taste Rating are more strongly correlated than pH and Taste Rating by 14%. An additional factor is the high degree of precision needed to measure TA. This raises the possibility that with a more precise data set there may be a stronger correlation than presents in the current analysis.

## Conclusion

TA and pH are both valuable quality metrics. TA has a stronger correlation with our Taste Rating and therefore may be a slightly better indicator than pH. Future investigation into the correlation between measures of Starter TA and Taste Rating of final product are another potential application of this metric.
